##**Statistical Inference Course Project Part1**##

*Haimo Liu, Sep.22th 2015*

###**Overview**###

In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda, where lambda is the rate parameter. I will investigate the distribution of averages of 40 exponentials within a thousand simulations.

```{r}
library(ggplot2)
```

###**Simulations**###

First we create a matrix with 1000 simulations, each simulation contains 40 samples of expotential distributions with lambda = 0.2. 

```{r}
sim = 1000
lambda = 0.2
n = 40
set.seed(10)
data = matrix(rexp(n*sim, lambda), sim);
```

Next, we calculate the mean of each simulation, and plot the distribution of the 1000 mean values

```{r}
mean_data = apply(data, 1, mean)
hist(mean_data)
```


###**Show the sample mean and compare it to the theoretical mean of the distribution**###

The theoretical mean of the distribution is **1/lambda = 5**. And sample mean is

```{r}
mean(mean_data)
```

###**Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution**###

The theoretical variance of the distribution is **1/((lambda)^2xn) = 0.625**. And the sample variance is:

```{r}
var(mean_data)
```


###**Show the distribution is approximately normal**###

First, let's look at the shape of the data:

```{r}
plot = data.frame(mean_data);
k = ggplot(plot, aes(x =mean_data)) 
k = k + geom_histogram(aes(y=..density..), colour="black",fill = "Blue")
k + geom_density(colour="Orange", size=2)
```

Next, let's compare the 95% confidence interval to that of a normal distribution

```{r}
con_interval_upper_bond = round (mean(mean_data) + 1.96*sd(mean_data)/sqrt(n),3)
theo_interval_upper_bond = 5 + 1.96*sqrt(0.625)/sqrt(n)
print(paste('Actual 95% confidence interval upper bond is', con_interval_upper_bond))
print(paste('Theoretical 95% confidence interval upper bond is', theo_interval_upper_bond))
```





